### 2020-04-18 reports/thesis/book.tex
Corrections

\begin{titlingpage}
\centering
{\LARGE Characterisation of Computed Tomography Noise in Projection Space with Applications to Additive Manufacturing \par}
\vspace{1cm}
{\Large Sherman Edea Lo\par}
{\Large Doctor of Philosophy in Statistics\par}
\vfill
\includegraphics[width = 0.75\textwidth]{../figures/frontCover.jpg}
\vfill
{\Large University of Warwick\par}
{\Large Department of Statistics\par}
{\Large September 2019\par}
\end{titlingpage}

@@

@@

The main statistical novelty was extending the empirical null \citep{efron2004large} to a filter where it can be applied to images. This has been implemented in \emph{ImageJ} \citep{abramoff2004image, schneider2012nih, perez2013image}. A by-product is the empirical null mean, which is effectively the mode filter, which can be used to smooth images.

@@

Voids in the order of millimetres were detectable using this method. Further work could include test samples made from materials such as titanium alloys (Ti-6Al-4V). This can be challenging because artefacts, such as streaks, can appear when using XCT for metal parts \citep{deman1998metal, nawaz2014metal}. Another problem is that because titanium products are created using a power-based method, the powder may become trapped \citep{brierley2018optimized} in designed voids. This results in voids appearing with less contrast in the projections, making it harder to detect. Different types of challenging defects could be investigated. For example, the orientation of cracks can affect the detectability of it \citep{wooldridge1997demonstrating}.

>>>>

\begin{titlingpage}
\centering
{\LARGE Characterisation of Computed Tomography Noise in Projection Space with Applications to Additive Manufacturing \par}
\vspace{1cm}
{\Large Sherman Edea Lo\par}
{\Large Doctor of Philosophy in Statistics\par}
\vfill
\includegraphics[width = 0.75\textwidth]{../figures/frontCover.jpg}
\vfill
{\Large University of Warwick\par}
{\Large Department of Statistics\par}
{\Large May 2020\par}
\end{titlingpage}

@@

This thesis is a contribution to real-time quality control in additive manufacturing.

@@

The main statistical novelty was extending the empirical null \citep{efron2004large} to a filter where it can be applied to images. This has been implemented in \emph{ImageJ} \citep{abramoff2004image, schneider2012nih, perez2013image}. A by-product is the empirical null mean, effectively the mode filter, which can be used to smooth images.

@@

Voids in the order of millimetres were detectable using this method. Further work could include test samples made from materials such as titanium alloys (Ti-6Al-4V). This can be challenging because artefacts, such as streaks, can appear when using XCT for metal parts \citep{deman1998metal, nawaz2014metal}. Another problem is that because titanium products are created using a power-based method, the powder may become trapped \citep{brierley2018optimized} in designed voids. This results in voids appearing with less contrast in the projections, making it harder to detect. Different types of challenging defects could be investigated, for example, the orientation of cracks can affect the detectability of it \citep{wooldridge1997demonstrating}.

### 2020-04-18 reports/thesis/chapters/3_compoundPoisson.tex
Corrections

\section{Conclusion}
The compound Poisson-gamma distribution was studied but it was found there were identifiability issues when fitting the model onto simulated data, in particular for high $\lambda$. It was found that the log-likelihood was quite flat because, by using numerous approximations, it was found that the determinant of the Hessian is zero. This cast doubts if a unique maximum likelihood estimator exists. Tighter bounds for the Hessian is beyond the scope of this thesis.

>>>>

\section{Conclusion}
The compound Poisson-gamma distribution was studied but it was found there were identifiability issues when fitting the model onto simulated data, in particular for high $\lambda$. It was found that the log-likelihood was quite flat because, by using numerous approximations, it was found that the determinant of the Hessian is about zero. This cast doubts if a unique maximum likelihood estimator exists. Tighter bounds for the Hessian is beyond the scope of this thesis.

### 2020-04-18 reports/thesis/chapters/4_variancePrediction.tex
Corrections
CpEmAlgorithm does not plot estimated std
roc calculates area exactly
Recoded NullIid and included biweighted mean
Corrections

The sample variance-mean data were obtained from the replicate projections. Let $x_{i,j}$ be the grey value of the $i$th pixel, ignoring any spatial information, from the $j$th replicate projection for $i=1,2,3,\ldots,N$ and $j=1,2,3,\ldots,m$. The sample mean and sample variance grey value for the $i$th pixel are
\begin{equation}
    \overline{x}_{i}=\frac{1}{m}\sum_{j=1}^m x_{i,j}
\end{equation}
and
\begin{equation}
    y_{i} =
    \frac{1}{m-1}
    \sum_{j=1}^m
        \left(
            x_{i,j} - \overline{x}_{i}
        \right)^2
\end{equation}
respectively. The symbol $y$ should not be confused with the latent varible in the previous chapter. Only pixels in the region of interest (ROI) were considered here, that is, pixels which represent the test sample. The ROI was created by manually segmenting the test sample from the projection.

@@

It was unusual to see an infection in the sample variance-mean data in \texttt{AbsNoFilter} at \ang{120} which was not captured by the GLM. A good explanation was not found but perhaps the absence of an x-ray filter may have contributed to this. As a result, the \texttt{AbsNoFilter} dataset was discarded in favour of the \texttt{AbsFilter} dataset from the next chapter.

There exist model selection methods such as lasso and elastic net \citep{tibshirani1996regression, zou2005regularization, friedman2010regularization} but they are catered for high dimensional problems involving hundreds of parameters. In this problem, only a few parameters were needed to find a good model. More flexible models such as non-parametric and machine learning methods may be used, however, they are slow and are unnecessary in a low dimensional problem with a large number of data points.

>>>>

The sample variance-mean data were obtained from the replicate projections. Let $x_{i,j}$ be the grey value of the $i$th pixel, ignoring any spatial information, from the $j$th replicate projection for $i=1,2,3,\ldots,N$ and $j=1,2,3,\ldots,m$. The sample mean and sample variance grey value for the $i$th pixel are
\begin{equation}
    x_{i}=\frac{1}{m}\sum_{j=1}^m x_{i,j}
\end{equation}
and
\begin{equation}
    y_{i} =
    \frac{1}{m-1}
    \sum_{j=1}^m
        \left(
            x_{i,j} - \overline{x}_{i}
        \right)^2
\end{equation}
respectively. The symbol $y$ should not be confused with the latent varible in the previous chapter. Only pixels in the region of interest (ROI) were considered here, that is, pixels which represent the test sample. The ROI was created by manually segmenting the test sample from the projection.

@@

It was unusual to see an inflection in the sample variance-mean data in \texttt{AbsNoFilter} at \ang{120} which was not captured by the GLM. A good explanation was not found but perhaps the absence of an x-ray filter may have contributed to this. As a result, the \texttt{AbsNoFilter} dataset was discarded in favour of the \texttt{AbsFilter} dataset from the next chapter.

There exist model selection methods such as lasso and elastic net \citep{tibshirani1996regression, zou2005regularization, friedman2010regularization} but they are catered for high dimensional problems involving hundreds of parameters. In this problem, only a few parameters were needed to find a good model thus forward stepwise selection was sufficient. More flexible models such as non-parametric and machine learning methods may be used, however, they are slow and are unnecessary in a low dimensional problem with a large number of data points.

### 2020-04-18 reports/thesis/chapters/5_inference.tex
Corrections

It is important to identity which quantites are random and which are not. For high photon rates, it was shown in the previous chapters that the grey values in the projection can be modelled using a Normal distribution. Thus, $\text{projection}_{x,y}$ is a random quantity. The simulation $\emph{aRTist}_{x,y}$ is not random because this was obtained through a computer software. Given the 19 selected projections used for training the variance-mean model, the variance prediction $\widehat{y}\left[\emph{aRTist}_{x,y}\right]$ is not random because the variance-mean model was fitted before encountering the remaining projection to be compared with \emph{aRTist}. The variance-mean model can be made random if a different set of 19 projections was used to train the model each time this inference was conducted, but this shall not be considered here.

@@

It can be noted that this mode finding method can be aplied to other distributions in general.

@@

The GLM fit in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} appeared to underfit, which suggest there may be a more complicated relationship between the optimal bandwidth and $n$, for example, a quadratic term can be added to the GLM fit. The method used in fitting the smoothing spline could contribute to the underfitting. For example, if the smoothing spline was not smooth enough so that multiple local minima appeared, it would add variance to the optimal bandwidth if the optimisation algorithm converges to a local minimum. A basic linear relationship was chosen to avoid overfitting and to keep the rule of thumb simple. It would also choose a large bandwidth for small $n$ which would improve the performance $\widehat{\mu}_0$ by working on a smoother density estimate for the price of a less optimal $s^2\left[\emph{aRTist}_{x,y}\right]$.

@@

The empirical null mean is effectively the mode and the empirical null standard deivation is a measure of dispersion. It was investigated how they compare with various estimates of central tendency and dispersion for varying sample sizes $n$. The empirical sampling distribution of these estimators was observed to look for properties such as bias, variance and robustness.

The central tendency estimates looked at were the median, the mean and Tukey's biweighted mean \citep{beaton1974fitting}. In summary, the biweighted mean weigh each data point using a smooth function of its residual and ignores any data points with residuals greater than a certain value. \emph{MATLAB}'s \texttt{robustfit} function was used for the biweighted mean and selects any tuning parmaeters.

The dispersion estimates looked at were the interquartile range (IQR) $\div 1.349$, the standard deviation and the median absolute deviation around the mode (MADA-mode) $\times 1.483$, . The normalisation constants were used so that MADA-mode and IQR can be compared directly to the standard deviation. MADA-mode was considered to investigate how estimators, which depends on the mode, perform.

@@

The empirical null standard deviation had some positive bias and it was found that this bias was quite sensitive to the bandwidth. Further fine-tuning on the bandwidth would be required for the empirical null standard deviation to be unbiased. For example, it appeared there was consistent positive bias for all $r$ investigated, perhaps suggesting the bandwidth parameter $a$ should be decreased ever so slightly. The source of bias could come from the fact that the bandwidth was tuned to minimise the log squared error rather than the squared bias.

It is interesting to compare the empirical null standard devaition with MADA-mode, in Figure \ref{fig:inference_NullIidStd}, because they both depend on the mode. The mode suffered from high variance with may affect dispersion estimators using the mode. The figure shows that MADA-mode has less variance than the empirical null standard deviation. This suggests that using the curvature of the density estimate at the mode is a main source of variance for the empirical null standard devation.

@@

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean1.eps}
        \caption{Mode}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean2.eps}
        \caption{Bisquare weighted mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean3.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean4.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd1.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd2.eps}
        \caption{MADA-mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd3.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd4.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidStd}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean1.eps}
        \caption{Mode}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean2.eps}
        \caption{Bisquared weighted mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean3.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean4.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMixtureMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd1.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd2.eps}
        \caption{MADA-mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd3.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd4.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidMixtureStd}
\end{figure}

>>>>

It is important to identify which quantities are random and which are not. For high photon rates, it was shown in the previous chapters that the grey values in the projection can be modelled using a Normal distribution. Thus, $\text{projection}_{x,y}$ is a random quantity. The simulation $\emph{aRTist}_{x,y}$ is not random because this was obtained through computer software. Given the 19 selected projections used for training the variance-mean model, the variance prediction $\widehat{y}\left[\emph{aRTist}_{x,y}\right]$ is not random because the variance-mean model was fitted before encountering the remaining projection to be compared with \emph{aRTist}. The variance-mean model can be made random if a different set of 19 projections was used to train the model each time this inference was conducted, but this shall not be considered here.

@@

It can be noted that this mode finding method can be applied to other distributions in general.

@@

The GLM fit in Figure \ref{fig:inference_BandwidthSelection2_ruleOfThumb} appeared to underfit, which suggest there may be a more complicated relationship between the optimal bandwidth and $n$, for example, a quadratic term can be added to the GLM fit. The method used in fitting the smoothing spline could contribute to the underfitting. For example, if the smoothing spline was not smooth enough so that multiple local minima appeared, it would add variance to the optimal bandwidth if the optimisation algorithm converges to a local minimum. A basic linear relationship was chosen to avoid overfitting and to keep the rule of thumb simple. It would also choose a large bandwidth for small $n$ which would improve the performance $\widehat{\mu}_0$ by working on a smoother density estimate for the price of a less optimal $\widehat{\sigma}_0$.

@@

The empirical null mean is effectively the mode and the empirical null standard deviation is a measure of dispersion. It was investigated how they compare with various estimates of central tendency and dispersion for varying sample sizes $n$. The empirical sampling distribution of these estimators was observed to look for properties such as bias, variance and robustness.

The central tendency estimates looked at were the median, the mean and Tukey's biweighted mean \citep{beaton1974fitting}. In summary, the biweighted mean weighs each data point using a smooth function of its residual and ignores any data points with residuals greater than a certain value. \emph{MATLAB}'s \texttt{robustfit} function was used for the biweighted mean and selects any tuning parameters.

The dispersion estimates looked at were the interquartile range (IQR) $\div 1.349$, the standard deviation and the median absolute deviation around the mode (MADA-mode) $\times 1.483$. The normalisation constants were used so that MADA-mode and IQR can be compared directly to the standard deviation. MADA-mode was considered to investigate how estimators, which depends on the mode, perform.

@@

The empirical null standard deviation had some positive bias and it was found that this bias was quite sensitive to the bandwidth. Further fine-tuning on the bandwidth would be required for the empirical null standard deviation to be unbiased. For example, it appeared there was a consistent positive bias for all $r$ investigated, perhaps suggesting the bandwidth parameter $a$ should be decreased ever so slightly. The source of bias could come from the fact that the bandwidth was tuned to minimise the log squared error rather than the squared bias.

It is interesting to compare the empirical null standard deviation with MADA-mode, in Figure \ref{fig:inference_NullIidStd}, because they both depend on the mode. The mode suffered from high variance which may affect dispersion estimators using the mode. The figure shows that MADA-mode has less variance than the empirical null standard deviation. This suggests that using the curvature of the density estimate at the mode is the main source of variance for the empirical null standard deviation.

@@

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean1.eps}
        \caption{Mode}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean2.eps}
        \caption{Bisquare weighted mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean4.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullMean3.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd1.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd2.eps}
        \caption{MADA-mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd4.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIid_nullStd3.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ standard Normal random variables. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidStd}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean1.eps}
        \caption{Mode}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean2.eps}
        \caption{Bisquared weighted mean}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean4.eps}
        \caption{Median}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullMean3.eps}
        \caption{Mean}
      \end{subfigure}
    }
  \caption{Different types of central tendency estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample mean.}
  \label{fig:inference_NullIidMixtureMean}
\end{figure}

\begin{figure}[p]
  \centering
    \centerline{
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd1.eps}
        \caption{Empirical null std}
      \end{subfigure}
      \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd2.eps}
        \caption{MADA-mode}
      \end{subfigure}
    }
    \centerline{
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd4.eps}
        \caption{IQR}
      \end{subfigure}
      \begin{subfigure}{\subSize}
        \includegraphics[width=\textwidth]{../figures/inference/NullIidMixture_nullStd3.eps}
        \caption{Standard deviation}
      \end{subfigure}
    }
  \caption{Different types of statistical dispersion estimators were used on $\pi r^2$ mixture of Gaussian random variables. The random variable has distribution $\normal(0,1)$ with probability $0.9$ and $\normal(3,1)$ with probability $0.1$. The boxplots represent the sampling distributions over 100 estimates. The dashed lines show the 95\% confidence interval for the sample variance.}
  \label{fig:inference_NullIidMixtureStd}
\end{figure}
